\textbf{Differences in COVID-19 vaccination in the province of Ontario across Health Regions and socio-economic strata}

### 2023-05-17 manuscript_v5/04-discussion.qmd
discussion updated

had significantly higher odds of vaccination than those in the Health Region of Toronto

In this study we hypothesized that differences in COVID-19 vaccination uptake were present between the Health Regions during between late 2021 and early 2022, aiming at determining which socio-demographic groups could be impacted by these disparities in order to provide decision-makers with information that could be used to develop policies focused on reducing or eliminating these differences and ensure that the Health Region model is able to fulfill its missing of improving health access for the inhabitants of Ontario.

Our results show differences in odds of vaccination across Ontario in certain socio-demographic groups. Specifically, those who identified as Arab/Middle Eastern, Black, Latin American, having mixed racial or ethnic background, or that belonged to other groups not explicitly included in the survey (Southeast Asian, Filipino, West Asian, and minority groups not identified elsewhere) had vaccination odds that were between a third and a half of that of individuals that identified as White/Caucasian (@fig-models). These results are consistent with previous studies that have shown lower vaccination rates in individuals with the same socio-demographic characteristics [@guay2022;@muhajarine2021;@carter2022;@hussain2022].

@@

However, the results also indicate that the place of habitation within the province for certain equity-deserving groups also affected the odds of vaccination (interaction terms of Health Region and Race, @fig-models). Specifically, this held true in the case of individuals identifying as Indigenous or with mixed racial background in the West Health Region, Black individuals in the Central Health Region, and Arab/Middle Eastern individuals in the East Health Region [@fig-models]. For these individuals, vaccination odds were lower when compared to the Toronto Health Region (Supplementary Figure A-6). Although it is likely that these findings have multiple causes, we indicate next some contributing factors that might help provide context in each case.

First, in the case of Indigenous, mixed, and Arab/Middle Eastern individuals is useful to disaggregate the data by LHINs, because most of the studies in the literature that have analyzed health in Ontario use the LHINs as the base of their analyses. Interestingly, for the first two groups most of the observations in the survey (for West Health Region) came from the Hamilton Niagara Haldimand Brant, South West, and Waterloo Wellington LHINs (**add table in Appendix**), whereas for Arab/Middle Eastern Individuals in the East Health Region the highest number of observations corresponded to the Champlain and Central East LHINs. Previous research has identified health disparities in these (mostly rural) regions, such as unequal distribution of primary care providers, increased mortality, and low pharmacist availability [@shah2019;@crighton2015;@timony2022].

Furthermore, there is an ongoing challenge for the health system of the province with regard to personalized healthcare for marginalized individuals. For example, the West Health Region has only two Aboriginal Health Access Centres in contrast to an Indigenous population of about 100,000 individuals [@ontariohealth]. Lack of access to personalized healthcare affects individuals that may mistrust the traditional healthcare system due to systemic racism or oppression, which is known to be the case for Indigenous and Black individuals in Canada where these rationales have been associated to lower vaccination among these groups [@smylie2022;@eissa2021]. Taken together, these aspects indicate ongoing multiple healthcare disparities that continue to affect equity-deserving individuals across the province.

The West Health Region is located in the leftmost southern part of the province, including the regions of Waterloo and Niagara, the counties of Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent. The Central Health Region includes the regions of Peel and Halton, the regional municipality of York, the district of Muskoka, the counties of Dufferin, Simcoe, Grey, and the city of Etobicoke.

There are some limitations to the present study. First, the data collection design, which allowed respondents to withdraw from the survey at any point, resulted in a high number of unique entries in the survey with multiple missing answers. Because we focused on entries that had complete observations in the covariates of interest for our analysis, it is possible that some information was not considered by excluding observations that had information in other variables (such as work from home, or number of persons in the household). However, we attempted to minimize this possibility by correcting the dataset using information from the Census. More granular corrections, which for example could be based on demographic information by municipality, could be used in the future to obtain a more accurate approximation to the population totals of the province. Moreover, our analysis did not consider the North West and North East Health Regions, due to the low number of entries from these areas in the survey (@fig-map). Although low representation from these areas is based on the fact that these regions only account for 5\% of the total population of Ontario, these regions are the home to more than 100,000 individuals that identify as Indigenous [@ontariohealth], a minority group that has historically suffered from reduced access to health care and discrimination [@mosby2021]. Therefore, there is a need for additional studies that focus on these low-populated Health Regions in Ontario where disparities in vaccination might be significant and understudied.

The results in this study are based on self-reported data, where bias might be present. However, because in the context of COVID-19 it has been shown that good agreement exists between self-reported and documented vaccination status [@stephenson2022], we believe that our data was able to provide a valid assessment of vaccination in the province. Finally, this study focused on first-dose vaccination status  within a relatively short time window, and therefore can only provide a snapshot of the societal dynamics behind the pandemic. Nonetheless, the results presented here can serve as a starting point to motivate the collection of robust longitudinal data that can be used to quantify geographical and temporal differences within vulnerable segments of the population, and that can be used to inform the development of adequate public health policies within the province of Ontario or across other provinces that aim to minimize disparities in health access.

>>>>

In this study we hypothesized that differences in COVID-19 vaccination uptake were present between the Health Regions during between late 2021 and early 2022, aiming at determining which socio-demographic groups could be impacted by these disparities in order to provide decision-makers with information that could be used to develop policies focused on reducing or eliminating these differences and ensure that the Health Region model is able to fulfill its mission of improving health access for the inhabitants of Ontario.

Our results show that indeed, there were differences in vaccination odds across Ontario in certain socio-demographic groups. Specifically, those who identified as Arab/Middle Eastern, Black, Latin American, having mixed racial or ethnic background, or that belonged to other groups not explicitly included in the survey (Southeast Asian, Filipino, West Asian, and minority groups not identified elsewhere) had vaccination odds that were between a third and a half of that of individuals that identified as White/Caucasian (@fig-models). These results are consistent with previous studies that have shown lower vaccination rates in individuals with the same socio-demographic characteristics [@guay2022;@muhajarine2021;@carter2022;@hussain2022].

@@

However, the results also indicate that the place of habitation affected the odds of vaccination for certain equity-deserving groups (interaction term of Health Region and Race, @fig-models). Specifically, this held true in the case of individuals identifying as Indigenous or with mixed racial background in the West Health Region, Black individuals in the Central Health Region, and Arab/Middle Eastern individuals in the East Health Region [@fig-models]. For these individuals, vaccination odds were lower when compared to the Toronto Health Region (Supplementary Figure A-6). We indicate next some contributing factors that might help provide context in each case.

First, it is useful to analyze the data using the LHINs, because most of the studies in the literature that have analyzed health in Ontario use the LHINs as the base of their analyses. Interestingly, for Indigenous and mixed individuals most of the observations in the survey for West Health Region came from the Hamilton Niagara Haldimand Brant, South West, and Waterloo Wellington LHINs (**add table in Appendix**), whereas for Arab/Middle Eastern Individuals in the East Health Region the highest number of observations corresponded to the Champlain and Central East LHINs. Previous research has identified health disparities in these (mostly rural) regions, such as unequal distribution of primary care providers, increased mortality, and low pharmacist availability [@shah2019;@crighton2015;@timony2022].

Furthermore, there is an ongoing challenge for the health system of the province with regard to personalized healthcare for marginalized individuals. For example, the West Health Region has only two Aboriginal Health Access Centres (community-led primary healthcare organizations focused on First Nations, Métis, and Inuit communities) to provide care to an estimated 100,000 Indigenous individuals living in the area [@ontariohealth]. Lack of access to personalized healthcare affects individuals that may mistrust the traditional healthcare system due to systemic racism or oppression, which is known to be the case for Indigenous and Black individuals in Canada. Indeed, these rationales have been associated to observed lower vaccination rates among these groups [@smylie2022;@eissa2021]. Taken together, this suggests that healthcare disparities that are specific to certain socio-demographic groups are associated with lower vaccination uptake.

There are some limitations to the present study. First, the data collection design, which allowed respondents to withdraw from the survey at any point, and that deployed the questions in a random manner resulted in an elevated number of missing observations without a definite pattern and complicated the implementation of sensitivity analyses. Therefore, we focused on entries that had complete answers, and corrected the data using population-wide information from the Census. However, more granular corrections would be needed to obtain more accurate estimates. For example, our analysis identified higher odds of vaccination in the Central and West Health Regions, but in this case these differences are likely to be driven by the proportion of White/Caucasian individuals, who had higher vaccination rates that other racial groups. Correcting for each racial/ethnic group in each Health Region can provide a more accurate estimation of region-wide vaccination rates. Unfortunately, at the moment this correction cannot be implemented as such stratification is not currently available in the Census data.

Additionally, our analysis did not consider the North West and North East Health Regions, due to the low number of entries from these areas in the survey (@fig-map). Low representation is expected as these regions only account for 5\% of the total population of Ontario, but in contrast, they have the highest proportion of Indigenous inhabitants [@ontariohealth]. In the context of personalized care, there is a need for collecting data that focuses on these Health Regions where additional health disparities might be present and possibly understudied.

The results in this study are based on self-reported data, where bias might be present. However, because in the context of COVID-19 it has been shown that good agreement exists between self-reported and documented vaccination status [@stephenson2022], we believe that our data was able to provide a valid assessment of vaccination in the province. Finally, although higher vaccination odds identified for January of 2022 in the model are in accordance with province-wide trends reported by Public Health Ontario (which show a 4\% increase between early December and January, in contrast to a 2.5\% increase between October and November [@ontario-covid]), these results are only a snapshot when compared to the overall duration of the pandemic.

Nonetheless, the results presented here can serve as a starting point to motivate the collection of robust longitudinal data that can be used to quantify geographical and temporal differences within vulnerable segments of the population, and that can be used to inform the development of adequate public health policies within the province of Ontario or across other provinces that aim to minimize disparities in health access.

### 2023-05-17 manuscript_v5/main.tex
discussion updated

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, boxrule=0pt, interior hidden, enhanced, breakable, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi

@@

had significantly higher odds of vaccination than those in the Health
Region of Toronto

In this study we hypothesized that differences in COVID-19 vaccination
uptake were present between the Health Regions during between late 2021
and early 2022, aiming at determining which socio-demographic groups
could be impacted by these disparities in order to provide
decision-makers with information that could be used to develop policies
focused on reducing or eliminating these differences and ensure that the
Health Region model is able to fulfill its missing of improving health
access for the inhabitants of Ontario.

Our results show differences in odds of vaccination across Ontario in
certain socio-demographic groups. Specifically, those who identified as
Arab/Middle Eastern, Black, Latin American, having mixed racial or
ethnic background, or that belonged to other groups not explicitly
included in the survey (Southeast Asian, Filipino, West Asian, and
minority groups not identified elsewhere) had vaccination odds that were
between a third and a half of that of individuals that identified as
White/Caucasian (Figure~\ref{fig-models}). These results are consistent
with previous studies that have shown lower vaccination rates in
individuals with the same socio-demographic
characteristics\textsuperscript{\protect\hyperlink{ref-guay2022}{18}--\protect\hyperlink{ref-hussain2022}{20},\protect\hyperlink{ref-carter2022}{48}}.

@@

However, the results also indicate that the place of habitation within
the province for certain equity-deserving groups also affected the odds
of vaccination (interaction terms of Health Region and Race,
Figure~\ref{fig-models}). Specifically, this held true in the case of
individuals identifying as Indigenous or with mixed racial background in
the West Health Region, Black individuals in the Central Health Region,
and Arab/Middle Eastern individuals in the East Health Region
Figure~\ref{fig-models}. For these individuals, vaccination odds were
lower when compared to the Toronto Health Region (Supplementary Figure
A-6). Although it is likely that these findings have multiple causes, we
indicate next some contributing factors that might help provide context
in each case.

First, in the case of Indigenous, mixed, and Arab/Middle Eastern
individuals is useful to disaggregate the data by LHINs, because most of
the studies in the literature that have analyzed health in Ontario use
the LHINs as the base of their analyses. Interestingly, for the first
two groups most of the observations in the survey (for West Health
Region) came from the Hamilton Niagara Haldimand Brant, South West, and
Waterloo Wellington LHINs (\textbf{add table in Appendix}), whereas for
Arab/Middle Eastern Individuals in the East Health Region the highest
number of observations corresponded to the Champlain and Central East
LHINs. Previous research has identified health disparities in these
(mostly rural) regions, such as unequal distribution of primary care
providers, increased mortality, and low pharmacist
availability\textsuperscript{\protect\hyperlink{ref-shah2019}{59}--\protect\hyperlink{ref-timony2022}{61}}.

Furthermore, there is an ongoing challenge for the health system of the
province with regard to personalized healthcare for marginalized
individuals. For example, the West Health Region has only two Aboriginal
Health Access Centres in contrast to an Indigenous population of about
100,000
individuals\textsuperscript{\protect\hyperlink{ref-ontariohealth}{62}}.
Lack of access to personalized healthcare affects individuals that may
mistrust the traditional healthcare system due to systemic racism or
oppression, which is known to be the case for Indigenous and Black
individuals in Canada where these rationales have been associated to
lower vaccination among these
groups\textsuperscript{\protect\hyperlink{ref-smylie2022}{63},\protect\hyperlink{ref-eissa2021}{64}}.
Taken together, these aspects indicate ongoing multiple healthcare
disparities that continue to affect equity-deserving individuals across
the province.

The West Health Region is located in the leftmost southern part of the
province, including the regions of Waterloo and Niagara, the counties of
Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand,
Brant, and Chatham-Kent. The Central Health Region includes the regions
of Peel and Halton, the regional municipality of York, the district of
Muskoka, the counties of Dufferin, Simcoe, Grey, and the city of
Etobicoke.

There are some limitations to the present study. First, the data
collection design, which allowed respondents to withdraw from the survey
at any point, resulted in a high number of unique entries in the survey
with multiple missing answers. Because we focused on entries that had
complete observations in the covariates of interest for our analysis, it
is possible that some information was not considered by excluding
observations that had information in other variables (such as work from
home, or number of persons in the household). However, we attempted to
minimize this possibility by correcting the dataset using information
from the Census. More granular corrections, which for example could be
based on demographic information by municipality, could be used in the
future to obtain a more accurate approximation to the population totals
of the province. Moreover, our analysis did not consider the North West
and North East Health Regions, due to the low number of entries from
these areas in the survey (Figure~\ref{fig-map}). Although low
representation from these areas is based on the fact that these regions
only account for 5\% of the total population of Ontario, these regions
are the home to more than 100,000 individuals that identify as
Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{62}}, a
minority group that has historically suffered from reduced access to
health care and
discrimination\textsuperscript{\protect\hyperlink{ref-mosby2021}{21}}.
Therefore, there is a need for additional studies that focus on these
low-populated Health Regions in Ontario where disparities in vaccination
might be significant and understudied.

It is also worth mentioning that province-wide vaccination rates for the
period of interest are somewhat different from those of the survey,
particularly in the case of those 55 years of age and older, which in
the survey had a vaccination rate of 72\%, against a rate of 88.4\%
reported for the closest age bracket (50 years of age and older)
reported by Public Health Ontario at the start of the period covered by
the data\textsuperscript{\protect\hyperlink{ref-ontario-covid}{65}}.
However, we found good agreement between the estimates from the model
and overall vaccination rates reported for Canada, which have been
relatively higher when compared to other high income
countries\textsuperscript{\protect\hyperlink{ref-dube2022}{66}}, and
with vaccination uptake rates across different age groups presented in
other
studies\textsuperscript{\protect\hyperlink{ref-guay2022}{18},\protect\hyperlink{ref-macdonald2021}{67}}.

In other words, although vaccination rates obtained from the survey were
slightly lower than the provincial estimates, these values still
represented a valid approximation to overall trends; this notion is
reinforced by the consistency in the proportion of vaccination rates
(Table~\ref{tbl-descriptive-stats}) and vaccination odds
(\textbf{?@fig-model-corr}) across the period covered by the survey,
which closely match the vaccination rates from Public Health Ontario and
which indicate that due to the relatively high coverage achieved in the
population at that point, vaccination rates increased by around 3\%
across all age groups during the three months covered by the
survey\textsuperscript{\protect\hyperlink{ref-ontario-covid}{65}}. It is
also important to mention that to this day, differences in vaccination
rates within the province continue. As of March of 2023 some regions
still have less than 75\% vaccination
rate\textsuperscript{\protect\hyperlink{ref-ontario-covid-map}{68}}, and
although data for the period analyzed in this study is not
publicly-available, it is likely that differences in vaccination rates
were higher at the time, being partially captured by the survey.

The results in this study are based on self-reported data, where bias
might be present. However, because in the context of COVID-19 it has
been shown that good agreement exists between self-reported and
documented vaccination
status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{69}}, we
believe that our data was able to provide a valid assessment of
vaccination in the province. Finally, this study focused on first-dose
vaccination status within a relatively short time window, and therefore
can only provide a snapshot of the societal dynamics behind the
pandemic. Nonetheless, the results presented here can serve as a
starting point to motivate the collection of robust longitudinal data
that can be used to quantify geographical and temporal differences
within vulnerable segments of the population, and that can be used to
inform the development of adequate public health policies within the
province of Ontario or across other provinces that aim to minimize
disparities in health access.

@@

\leavevmode\vadjust pre{\hypertarget{ref-ontario-covid}{}}%
\CSLLeftMargin{65. }%
\CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed February 27,
2023.
\url{https://www.publichealthontario.ca/en/data-and-analysis/infectious-disease/covid-19-data-surveillance/covid-19-data-tool?tab=vaccine}}

\leavevmode\vadjust pre{\hypertarget{ref-dube2022}{}}%
\CSLLeftMargin{66. }%
\CSLRightInline{Dubé E, Gagnon D, MacDonald N. Between persuasion and
compulsion: The case of {COVID}-19 vaccination in canada.
\emph{Vaccine}. 2022;40(29):3923-3926.
doi:\href{https://doi.org/10.1016/j.vaccine.2022.05.053}{10.1016/j.vaccine.2022.05.053}}

\leavevmode\vadjust pre{\hypertarget{ref-macdonald2021}{}}%
\CSLLeftMargin{67. }%
\CSLRightInline{MacDonald NE, Comeau J, Dubé Ève, et al. Royal society
of canada {COVID}-19 report: Enhancing {COVID}-19 vaccine acceptance in
canada. Blais JM, ed. \emph{{FACETS}}. 2021;6:1184-1246.
doi:\href{https://doi.org/10.1139/facets-2021-0037}{10.1139/facets-2021-0037}}

\leavevmode\vadjust pre{\hypertarget{ref-ontario-covid-map}{}}%
\CSLLeftMargin{68. }%
\CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed March 24, 2023.
\url{https://www.publichealthontario.ca/en/data-and-analysis/infectious-disease/covid-19-data-surveillance/covid-19-data-tool?tab=maps}}

\leavevmode\vadjust pre{\hypertarget{ref-stephenson2022}{}}%
\CSLLeftMargin{69. }%
\CSLRightInline{Stephenson M, Olson SM, Self WH, et al. Ascertainment of
vaccination status by self-report versus source documentation: Impact on
measuring {COVID}-19 vaccine effectiveness. \emph{Influenza and Other
Respiratory Viruses}. 2022;16(6):1101-1111.
doi:\href{https://doi.org/10.1111/irv.13023}{10.1111/irv.13023}}

>>>>

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, boxrule=0pt, sharp corners, interior hidden, enhanced]}{\end{tcolorbox}}\fi

@@

In this study we hypothesized that differences in COVID-19 vaccination
uptake were present between the Health Regions during between late 2021
and early 2022, aiming at determining which socio-demographic groups
could be impacted by these disparities in order to provide
decision-makers with information that could be used to develop policies
focused on reducing or eliminating these differences and ensure that the
Health Region model is able to fulfill its mission of improving health
access for the inhabitants of Ontario.

Our results show that indeed, there were differences in vaccination odds
across Ontario in certain socio-demographic groups. Specifically, those
who identified as Arab/Middle Eastern, Black, Latin American, having
mixed racial or ethnic background, or that belonged to other groups not
explicitly included in the survey (Southeast Asian, Filipino, West
Asian, and minority groups not identified elsewhere) had vaccination
odds that were between a third and a half of that of individuals that
identified as White/Caucasian (Figure~\ref{fig-models}). These results
are consistent with previous studies that have shown lower vaccination
rates in individuals with the same socio-demographic
characteristics\textsuperscript{\protect\hyperlink{ref-guay2022}{18}--\protect\hyperlink{ref-hussain2022}{20},\protect\hyperlink{ref-carter2022}{48}}.

@@

However, the results also indicate that the place of habitation affected
the odds of vaccination for certain equity-deserving groups (interaction
term of Health Region and Race, Figure~\ref{fig-models}). Specifically,
this held true in the case of individuals identifying as Indigenous or
with mixed racial background in the West Health Region, Black
individuals in the Central Health Region, and Arab/Middle Eastern
individuals in the East Health Region Figure~\ref{fig-models}. For these
individuals, vaccination odds were lower when compared to the Toronto
Health Region (Supplementary Figure A-6). We indicate next some
contributing factors that might help provide context in each case.

First, it is useful to analyze the data using the LHINs, because most of
the studies in the literature that have analyzed health in Ontario use
the LHINs as the base of their analyses. Interestingly, for Indigenous
and mixed individuals most of the observations in the survey for West
Health Region came from the Hamilton Niagara Haldimand Brant, South
West, and Waterloo Wellington LHINs (\textbf{add table in Appendix}),
whereas for Arab/Middle Eastern Individuals in the East Health Region
the highest number of observations corresponded to the Champlain and
Central East LHINs. Previous research has identified health disparities
in these (mostly rural) regions, such as unequal distribution of primary
care providers, increased mortality, and low pharmacist
availability\textsuperscript{\protect\hyperlink{ref-shah2019}{59}--\protect\hyperlink{ref-timony2022}{61}}.

Furthermore, there is an ongoing challenge for the health system of the
province with regard to personalized healthcare for marginalized
individuals. For example, the West Health Region has only two Aboriginal
Health Access Centres (community-led primary healthcare organizations
focused on First Nations, Métis, and Inuit communities) to provide care
to an estimated 100,000 Indigenous individuals living in the
area\textsuperscript{\protect\hyperlink{ref-ontariohealth}{62}}. Lack of
access to personalized healthcare affects individuals that may mistrust
the traditional healthcare system due to systemic racism or oppression,
which is known to be the case for Indigenous and Black individuals in
Canada. Indeed, these rationales have been associated to observed lower
vaccination rates among these
groups\textsuperscript{\protect\hyperlink{ref-smylie2022}{63},\protect\hyperlink{ref-eissa2021}{64}}.
Taken together, this suggests that healthcare disparities that are
specific to certain socio-demographic groups are associated with lower
vaccination uptake.

There are some limitations to the present study. First, the data
collection design, which allowed respondents to withdraw from the survey
at any point, and that deployed the questions in a random manner
resulted in an elevated number of missing observations without a
definite pattern and complicated the implementation of sensitivity
analyses. Therefore, we focused on entries that had complete answers,
and corrected the data using population-wide information from the
Census. However, more granular corrections would be needed to obtain
more accurate estimates. For example, our analysis identified higher
odds of vaccination in the Central and West Health Regions, but in this
case these differences are likely to be driven by the proportion of
White/Caucasian individuals, who had higher vaccination rates that other
racial groups. Correcting for each racial/ethnic group in each Health
Region can provide a more accurate estimation of region-wide vaccination
rates. Unfortunately, at the moment this correction cannot be
implemented as such stratification is not currently available in the
Census data.

Additionally, our analysis did not consider the North West and North
East Health Regions, due to the low number of entries from these areas
in the survey (Figure~\ref{fig-map}). Low representation is expected as
these regions only account for 5\% of the total population of Ontario,
but in contrast, they have the highest proportion of Indigenous
inhabitants\textsuperscript{\protect\hyperlink{ref-ontariohealth}{62}}.
In the context of personalized care, there is a need for collecting data
that focuses on these Health Regions where additional health disparities
might be present and possibly understudied.

The results in this study are based on self-reported data, where bias
might be present. However, because in the context of COVID-19 it has
been shown that good agreement exists between self-reported and
documented vaccination
status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{65}}, we
believe that our data was able to provide a valid assessment of
vaccination in the province. Finally, although higher vaccination odds
identified for January of 2022 in the model are in accordance with
province-wide trends reported by Public Health Ontario (which show a 4\%
increase between early December and January, in contrast to a 2.5\%
increase between October and
November\textsuperscript{\protect\hyperlink{ref-ontario-covid}{66}}),
these results are only a snapshot when compared to the overall duration
of the pandemic.

Nonetheless, the results presented here can serve as a starting point to
motivate the collection of robust longitudinal data that can be used to
quantify geographical and temporal differences within vulnerable
segments of the population, and that can be used to inform the
development of adequate public health policies within the province of
Ontario or across other provinces that aim to minimize disparities in
health access.

@@

\leavevmode\vadjust pre{\hypertarget{ref-stephenson2022}{}}%
\CSLLeftMargin{65. }%
\CSLRightInline{Stephenson M, Olson SM, Self WH, et al. Ascertainment of
vaccination status by self-report versus source documentation: Impact on
measuring {COVID}-19 vaccine effectiveness. \emph{Influenza and Other
Respiratory Viruses}. 2022;16(6):1101-1111.
doi:\href{https://doi.org/10.1111/irv.13023}{10.1111/irv.13023}}

\leavevmode\vadjust pre{\hypertarget{ref-ontario-covid}{}}%
\CSLLeftMargin{66. }%
\CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed February 27,
2023.
\url{https://www.publichealthontario.ca/en/data-and-analysis/infectious-disease/covid-19-data-surveillance/covid-19-data-tool?tab=vaccine}}
